A Logistic Regression Approach to CoIL Challenge 2000

### 2019-12-11 final project/Manuscript.Rmd
Cleaned up Manuscript

---
title             : "A Logistic Regression Approach to CoIL Challenge 2000"
shorttitle        : " "
author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
abstract: |
  A logistic regression based solution to the CoIL Challenge 2000 is described. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
  <!-- https://tinyurl.com/ybremelq 
    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  -->
keywords          : "CoIL Challenge, Logistic Regression"
wordcount         : "X"
bibliography      : ["Manuscript.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output: 
  word_document:
    reference_docx: word-template.docx
  papaja::apa6_pdf:
    #citation_package: natbib #biblatex
  html_document: default
---

```{r setup, include = FALSE}
# Documentation at https://crsh.github.io/papaja_man/
# If you have no use for TeX beyond rendering R Markdown documents, I recommend you use TinyTex.
# install.packages("tinytex")
# tinytex::install_tinytex()
if (!"papaja" %in% rownames(installed.packages())) {
  devtools::install_github("crsh/papaja")
}
library("papaja")
library(kableExtra)
library(tidyverse)
library(pander)

# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set(output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if (is.na(comment)) comment <- ""
  can_null <- grepl(paste0(comment, "\\s*\\[\\d?\\]"),
                     x, perl = TRUE)
  do_null <- isTRUE(knitr::opts_current$get("null_prefix"))
  if (can_null && do_null) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr("\\]", x)[1] - 1
    # Two cases: start or newline
    re <- paste0("^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub(re, rep,  x)
    re <- paste0("\\\n.{", align_index, "}\\]")
    rep <- paste0("\n", comment)
    x <- gsub(re, rep,  x)
  }

  default_output_hook(x, options)

})
knitr::opts_template$set("kill_prefix" = list(comment = NA, null_prefix = TRUE))
```

```{r analysis-preferences, echo=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r data, echo=FALSE}
prepare_data <- function(df){
  names(df) <- c("MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD",
               "MGODRK", "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA",
               "MRELOV", "MFALLEEN", "MFGEKIND", "MFWEKIND", "MOPLHOOG",
               "MOPLMIDD", "MOPLLAAG", "MBERHOOG", "MBERZELF", "MBERBOER",
               "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA", "MSKB1", "MSKB2",
               "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2", "MAUT0",
               "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
               "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART",
               "PWABEDR", "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO",
               "PVRAAUT",  "PAANHANG", "PTRACTOR", "PWERKT", "PBROM", "PLEVEN",
               "PPERSONG", "PGEZONG", "PWAOREG", "PBRAND", "PZEILPL",
               "PPLEZIER", "PFIETS", "PINBOED", "PBYSTAND", "AWAPART",
               "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT", "AMOTSCO",
               "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM", "ALEVEN",
               "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
               "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")

  MOSTYPE_labels <- c("1" = "1 High Income, expensive child",
                      "2" = "2 Very Important Provincials",
                      "3" = "3 High status seniors",
                      "4" = "4 Affluent senior apartments",
                      "5" = "5 Mixed seniors",
                      "6" = "6 Career and childcare",
                      "7" = "7 Dinki's (double income no kids)",
                      "8" = "8 Middle class families",
                      "9" = "9 Modern, complete families",
                      "10" = "10 Stable family",
                      "11" = "11 Family starters",
                      "12" = "12 Affluent young families",
                      "13" = "13 Young all american family",
                      "14" = "14 Junior cosmopolitan",
                      "15" = "15 Senior cosmopolitans",
                      "16" = "16 Students in apartments",
                      "17" = "17 Fresh masters in the city",
                      "18" = "18 Single youth",
                      "19" = "19 Suburban youth",
                      "20" = "20 Ethnically diverse",
                      "21" = "21 Young urban have-nots",
                      "22" = "22 Mixed apartment dwellers",
                      "23" = "23 Young and rising",
                      "24" = "24 Young, low educated", 
                      "25" = "25 Young seniors in the city",
                      "26" = "26 Own home elderly",
                      "27" = "27 Seniors in apartments",
                      "28" = "28 Residential elderly",
                      "29" = "29 Porchless seniors: no front yard",
                      "30" = "30 Religious elderly singles",
                      "31" = "31 Low income catholics",
                      "32" = "32 Mixed seniors",
                      "33" = "33 Lower class large families",
                      "34" = "34 Large family, employed child",
                      "35" = "35 Village families",
                      "36" = "36 Couples with teens 'Married with children'",
                      "37" = "37 Mixed small town dwellers",
                      "38" = "38 Traditional families",
                      "39" = "39 Large religous families",
                      "40" = "40 Large family farms",
                      "41" = "41 Mixed rurals")

MGEMLEEF_labels <- c("1" = "20-30 years",
                     "2" = "30-40 years",
                     "3" = "40-50 years",
                     "4" = "50-60 years",
                     "5" = "60-70 years",
                     "6" = "70-80 years")

MOSHOOFD_labels <- c("1" = "Successful hedonists",
                     "2" = "Driven Growers",
                     "3" = "Average Family",
                     "4" = "Career Loners",
                     "5" = "Living well",
                     "6" = "Cruising Seniors",
                     "7" = "Retired and Religeous",
                     "8" = "Family with grown ups",
                     "9" = "Conservative families",
                     "10" = "Farmers")

MGODRK_labels <- c("0" = "0%",
                   "1" = "1 - 10%",
                   "2" = "11 - 23%",
                   "3" = "24 - 36%",
                   "4" = "37 - 49%",
                   "5" = "50 - 62%",
                   "6" = "63 - 75%",
                   "7" = "76 - 88%",
                   "8" = "89 - 99%",
                   "9" = "100%")

PWAPART_labels <- c("0" = "f 0",
                    "1" = "f 1 – 49",
                    "2" = "f 50 – 99",
                    "3" = "f 100 – 199",
                    "4" = "f 200 – 499",
                    "5" = "f 500 – 999",
                    "6" = "f 1000 – 4999",
                    "7" = "f 5000 – 9999",
                    "8" = "f 10.000 - 19.999",
                    "9" = "f 20.000 - ?")

df %>%
  mutate(MOSTYPE = as.factor(MOSTYPE),
         MGEMLEEF = as.factor(MGEMLEEF),
         MOSHOOFD = as.factor(MOSHOOFD),
         MGODRK = as.factor(MGODRK),
         PWAPART = as.factor(PWAPART),
         CARAVAN = as.factor(CARAVAN)) %>%
  mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
         MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
         MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
         MGODRK = recode(MGODRK, !!!MGODRK_labels),
         PWAPART = recode(PWAPART, !!!PWAPART_labels))
}

eval <- read.delim("data/ticeval2000.txt", header = FALSE)
temp <- read.delim("data/tictgts2000.txt", header = FALSE)
eval$CARAVAN <- temp$V1
eval <- prepare_data(eval)
df <- prepare_data(read.delim("data/ticdata2000.txt", header = FALSE))
```
# Introduction
<!-- Describe the background and motivation of your problem.-->

Businesses use data science to extract insights from data.  It has many practical business applications.  Identifying households to include in a marketing campaign is one application.  One example using real world data is the Computational Intelligence and Learning (CoIL) Challenge. The CoIL Challenge competition was held from March 17 to May 8 in 2000.  The challenge is to:

1.  Identify potential customers for an insurance policy; and

2.  Provide a description of this customer base.

```{r, echo=FALSE}
# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83,
                                          86, 91, 94, 95, 96, 97, 99, 101, 103,
                                          104, 105, 106, 107, 108, 109, 110,
                                          111, 112, 115, 121),
                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3,
                                      1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1,
                                      1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Number of Correct Predictions") +
  ylab("CoIL Challenge Submissions") +
  xlim(37, 122)
```

In total 147 participants registered and 43 submitted solutions [@Putten]. The maximum number of policyowners that could be found was 238.  The submissions identified 95 policy owners on average.  The winning model [@Elkan] identified 121 policy owners.  Random selection results in identifying 42 policy owners. The standard benchmark tests result in 94 (k-nearest neighbor), 102 (naïve bayes), 105 (neural networks) and 118 (linear) policy owners. [@Putten].  In this paper we set out to complete the first part of the COIL Challenge. <!--**SUMARISE FINDINGS?**-->

@@

Participants used a variety of approaches in formulating their submissions including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].

<!--Our investigation is similar to @Doornik in that we used a LOGIT model. **Reproduce his model and describe the results**-->

@@

<!--Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

# Experimentation and Results

<!--Describe the specifics of what you did (data exploration, data preparation, model building, model selection, model evaluation, etc.), and what you found out (statistical analyses, interpretation and discussion of the results, etc.).

We used `r cite_r("Manuscript.bib")` for all our analyses.
-->

# Discussion and Conclusions

<!--Conclude your findings, limitations, and suggest areas for future work.-->

# References
```{r create_references, echo=FALSE}
r_refs(file = "Manuscript.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

# Appendices

<!--
- Supplemental tables and/or figures.
-->

@@

>>>>

---
title: "A Logistic Regression Approach to CoIL Challenge 2000"
author: "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
bibliography: "Manuscript.bib"
output: 
  word_document:
    reference_docx: word-template.docx
---

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(pander)

prepare_data <- function(df){
  names(df) <- c(
    "MOSTYPE", "MAANTHUI", "MGEMOMV", "MGEMLEEF", "MOSHOOFD", "MGODRK",
    "MGODPR", "MGODOV", "MGODGE", "MRELGE", "MRELSA", "MRELOV", "MFALLEEN",
    "MFGEKIND", "MFWEKIND", "MOPLHOOG", "MOPLMIDD", "MOPLLAAG", "MBERHOOG",
    "MBERZELF", "MBERBOER", "MBERMIDD", "MBERARBG", "MBERARBO", "MSKA",
    "MSKB1", "MSKB2", "MSKC", "MSKD", "MHHUUR", "MHKOOP", "MAUT1", "MAUT2",
    "MAUT0", "MZFONDS", "MZPART", "MINKM30", "MINK3045", "MINK4575",
    "MINK7512", "MINK123M", "MINKGEM", "MKOOPKLA", "PWAPART", "PWABEDR", 
    "PWALAND", "PPERSAUT", "PBESAUT", "PMOTSCO", "PVRAAUT",  "PAANHANG",
    "PTRACTOR", "PWERKT", "PBROM", "PLEVEN", "PPERSONG", "PGEZONG",
    "PWAOREG", "PBRAND", "PZEILPL", "PPLEZIER", "PFIETS", "PINBOED",
    "PBYSTAND", "AWAPART", "AWABEDR", "AWALAND", "APERSAUT", "ABESAUT",
    "AMOTSCO", "AVRAAUT", "AAANHANG", "ATRACTOR", "AWERKT", "ABROM",
    "ALEVEN", "APERSONG", "AGEZONG", "AWAOREG",  "ABRAND", "AZEILPL",
    "APLEZIER", "AFIETS", "AINBOED", "ABYSTAND", "CARAVAN")

  MOSTYPE_labels <- c(
    "1" = "1 High Income, expensive child",
    "2" = "2 Very Important Provincials",
    "3" = "3 High status seniors",
    "4" = "4 Affluent senior apartments",
    "5" = "5 Mixed seniors",
    "6" = "6 Career and childcare",
    "7" = "7 Dinki's (double income no kids)",
    "8" = "8 Middle class families",
    "9" = "9 Modern, complete families",
    "10" = "10 Stable family",
    "11" = "11 Family starters",
    "12" = "12 Affluent young families",
    "13" = "13 Young all american family",
    "14" = "14 Junior cosmopolitan",
    "15" = "15 Senior cosmopolitans",
    "16" = "16 Students in apartments",
    "17" = "17 Fresh masters in the city",
    "18" = "18 Single youth",
    "19" = "19 Suburban youth",
    "20" = "20 Ethnically diverse",
    "21" = "21 Young urban have-nots",
    "22" = "22 Mixed apartment dwellers",
    "23" = "23 Young and rising",
    "24" = "24 Young, low educated",
    "25" = "25 Young seniors in the city",
    "26" = "26 Own home elderly",
    "27" = "27 Seniors in apartments",
    "28" = "28 Residential elderly",
    "29" = "29 Porchless seniors: no front yard",
    "30" = "30 Religious elderly singles",
    "31" = "31 Low income catholics",
    "32" = "32 Mixed seniors",
    "33" = "33 Lower class large families",
    "34" = "34 Large family, employed child",
    "35" = "35 Village families",
    "36" = "36 Couples with teens 'Married with children'",
    "37" = "37 Mixed small town dwellers",
    "38" = "38 Traditional families",
    "39" = "39 Large religous families",
    "40" = "40 Large family farms",
    "41" = "41 Mixed rurals")

  MGEMLEEF_labels <- c(
    "1" = "20-30 years",
    "2" = "30-40 years",
    "3" = "40-50 years",
    "4" = "50-60 years",
    "5" = "60-70 years",
    "6" = "70-80 years")

  MOSHOOFD_labels <- c(
    "1" = "Successful hedonists",
    "2" = "Driven Growers",
    "3" = "Average Family",
    "4" = "Career Loners",
    "5" = "Living well",
    "6" = "Cruising Seniors",
    "7" = "Retired and Religeous",
    "8" = "Family with grown ups",
    "9" = "Conservative families",
    "10" = "Farmers")

  MGODRK_labels <- c(
    "0" = "0%",
    "1" = "1 - 10%",
    "2" = "11 - 23%",
    "3" = "24 - 36%",
    "4" = "37 - 49%",
    "5" = "50 - 62%",
    "6" = "63 - 75%",
    "7" = "76 - 88%",
    "8" = "89 - 99%",
    "9" = "100%")

  PWAPART_labels <- c(
    "0" = "f 0",
    "1" = "f 1 - 49",
    "2" = "f 50 - 99",
    "3" = "f 100 - 199",
    "4" = "f 200 - 499",
    "5" = "f 500 - 999",
    "6" = "f 1000 - 4999",
    "7" = "f 5000 - 9999",
    "8" = "f 10,000 - 19,999",
    "9" = "f 20,000 - ?")

  set_to_1 <- c(12, 8, 6, 7, 1, 2, 36, 3, 37, 10, 20, 38, 11)

  df %>%
    mutate(CUSTOMER_TYPE = ifelse(MOSTYPE %in% set_to_1, 1, 0)) %>%
    mutate(CUSTOMER_TYPE = as.factor(CUSTOMER_TYPE)) %>%
    mutate(MOSTYPE = as.factor(MOSTYPE),
           MGEMLEEF = as.factor(MGEMLEEF),
           MOSHOOFD = as.factor(MOSHOOFD),
           MGODRK = as.factor(MGODRK),
           PWAPART = as.factor(PWAPART),
           CARAVAN = as.factor(CARAVAN)) %>%
    mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
           MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
           MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
           MGODRK = recode(MGODRK, !!!MGODRK_labels),
           PWAPART = recode(PWAPART, !!!PWAPART_labels))
}

eval <- read.delim("data/ticeval2000.txt", header = FALSE)
temp <- read.delim("data/tictgts2000.txt", header = FALSE)
eval$CARAVAN <- temp$V1
eval <- prepare_data(eval)
df <- prepare_data(read.delim("data/ticdata2000.txt", header = FALSE))
```

# Abstract

A logistic regression based solution to the CoIL Challenge 2000 is described. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.

*Key words: CoIL Challenge, Logistic Regression*

# Introduction

Businesses use data science to extract insights from data.  One pratical application is identifying households to include in a marketing campaign.  In this paper we set out to identify potential customers for an insurance product using real world data from the Computational Intelligence and Learning (CoIL) Challenge.

@@

Fourty-three other research teams have attempted to identify potential caravan insurance customers [@Putten].  They used a variety of approaches including: Boosted Decision Tree [@McKone], Classification and Regression Tree (CART) [@Simmonds], Classification Trees with Bagging [@White], C4.5 [@Rickets; @Seewald], Evolutionary Algorithm [@Koudijs], Fuzzy Classifier [@Abonyi; @Kaymak], Genetic Algorithms and Hill-climbers [@Carter], Inductive Learning by Logic Minimization (ILLM) [@Gamberger; @Smuc], Instance Based Reasoning (iBARET) [@Miksovsky], K-Means [@Vesanto], KXEN [@Bera], LOGIT [@Doornik], Mask Perceptron with Boosting [@Leckie], Midos Algorithm [@Krogel], N-Tuple Classifier [@Jorgensen], Naïve Bayes [@Elkan; @Kontkanen], Neural Networks[@Brierley; @Crocoll; @Kim; @Shtovba_Mashnitskiy], Phase Intervals and Genetic Algorithms [@Shtovba], Scoring System [@Lewandowski], Support Vector Machines[@Keerthi], and XCS [@Greenyer].

```{r, echo=FALSE}
# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121), n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity", fill = "#377eb8") +
  xlab("Number of Correct Predictions") +
  ylab("CoIL Challenge Submissions") +
  xlim(37, 122) +
  scale_fill_brewer(palette = "Set1")
```

The maximum number of potential policy owners that could be found is 238.  Previous researchers identified 95 policy owners on average.  The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.  A meta analysis found that simpler algorithms tended to outpreform more complicated ones [@Putten].

@@

<!-- Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection. -->

# Appendix

@@

## References
